---
title: "Sepsis Prevalence and Mortality Risk in ICU Patients"
subtitle: "A MIMIC-III Database Analysis"
author: "Ernest Ceballos, Júlia Galimany, Oriol Galimany"
institute: "Universitat Rovira i Virgili"
date: "January 2026"
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    slide_level: 2
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Introduction

## What is Sepsis?

- **Definition**: Life-threatening organ dysfunction caused by dysregulated host response to infection

- **Clinical Impact**:
  - Leading cause of mortality in ICUs worldwide
  - Rapid progression requires early identification
  - Affects millions globally each year

- **Challenge**: Can we predict who is at highest risk?

## Research Objectives

1. **Estimate sepsis prevalence** among ICU patients

2. **Compare mortality rates** between sepsis and non-sepsis patients

3. **Develop a predictive model** for mortality risk in sepsis patients

4. **Identify key risk factors** associated with mortality

# Methods

## Data Source: MIMIC-III

- **Database**: Medical Information Mart for Intensive Care III (v1.4)
- **Institution**: Beth Israel Deaconess Medical Center
- **Period**: 2001-2012
- **Content**: De-identified health data from >40,000 ICU admissions

## Cohort Selection

**Inclusion Criteria:**

- Adult patients (age ≥ 16 years)
- At least one ICU admission recorded

**Final Cohort:** 61,524 unique ICU admissions

**Sepsis Identification (ICD-9 codes):**

| Code | Description |
|------|-------------|
| 038.x | Septicemia |
| 99591 | Sepsis |
| 99592 | Severe sepsis |
| 78552 | Septic shock |

## Statistical Analysis

- **Prevalence**: Proportion of ICU stays with sepsis diagnosis
- **Mortality comparison**: Chi-square test
- **Predictive modeling**: Logistic regression

**Model predictors:**

- Age, Gender, Ethnicity
- Insurance type
- ICD-9 diagnosis code

# Results

## Sepsis Prevalence

\begin{center}
\textbf{8.24\% of ICU admissions had sepsis}
\end{center}

- **Total ICU admissions**: 61,524
- **Sepsis cases**: 5,071

\vspace{0.5cm}

**Approximately 1 in 12 ICU patients has sepsis**

## Prevalence by Demographics

### By Gender:
- **Males**: 8.81% (higher)
- **Females**: 7.51%

### By Age Group:
| Age Group | Prevalence |
|-----------|------------|
| 16-40 | 6.4% |
| 40-60 | 8.2% |
| 60-80 | 8.7% |
| 80+ | 8.1% |

*Prevalence increases with age until 60-80 years*

## Mortality: Sepsis vs Non-Sepsis

```{r mortality-comparison, echo=FALSE}
# Create data for visualization
mortality_data <- data.frame(
  Group = c("With Sepsis", "Without Sepsis"),
  Mortality = c(24.39, 9.93)
)
```

| Group | Mortality Rate |
|-------|---------------|
| **With Sepsis** | **24.39%** |
| Without Sepsis | 9.93% |

\vspace{0.5cm}

\begin{center}
\textbf{Sepsis patients have 2.5× higher mortality risk}
\end{center}

*p < 0.001 (Chi-square test)*

## Predictive Model: ROC Curve

**Logistic Regression Performance:**

| Metric | Value |
|--------|-------|
| AUC | 0.622 |
| Optimal Threshold | 0.334 |
| Accuracy | 58.3% |
| Sensitivity | 60.4% |
| Specificity | 57.3% |
| PPV | 40.6% |
| NPV | 74.9% |

## Model Interpretation

**Key Findings:**

- AUC of 0.622 indicates **modest discriminative ability**
- Model performs better than random (AUC > 0.5)
- **Demographic factors alone are insufficient** for accurate risk stratification

**Significant predictors:**

- Age (older → higher risk)
- Insurance type (proxy for socioeconomic status)
- Specific ICD-9 codes

# Discussion

## Key Insights

1. **High Disease Burden**
   - 1 in 12 ICU patients affected
   - 2.5× mortality risk

2. **Demographic Patterns**
   - Higher prevalence in males
   - Increasing prevalence with age

3. **Model Limitations**
   - Demographic predictors insufficient (AUC = 0.622)
   - Need for clinical variables (lactate, vitals, etc.)

## Limitations

- **Retrospective design**: Cannot establish causality
- **Single center**: Results may not generalize
- **ICD-9 coding**: Possible under/over-reporting
- **Independence assumption**: Same patient may have multiple admissions
- **Missing clinical variables**: Vital signs, lab values excluded

## Clinical Implications

- Sepsis screening should be **universal in ICU settings**
- **Demographic factors alone cannot guide** treatment decisions
- Future models should incorporate:
  - Vital signs (temperature, heart rate, BP)
  - Laboratory values (lactate, WBC, creatinine)
  - Organ dysfunction scores (SOFA, qSOFA)

# Conclusion

## Summary

\begin{block}{Main Findings}
\begin{enumerate}
\item Sepsis prevalence: \textbf{8.24\%} of ICU admissions
\item Mortality risk: \textbf{2.5× higher} with sepsis
\item Predictive model: \textbf{AUC = 0.622} (modest performance)
\end{enumerate}
\end{block}

\vspace{0.5cm}

**Take-home message:** 

Sepsis remains a critical clinical challenge. Effective risk stratification requires integration of clinical parameters beyond basic demographics.

## Future Directions

1. **Incorporate clinical variables**
   - Vital signs, laboratory values
   - Organ dysfunction scores

2. **Machine learning approaches**
   - Random forests, neural networks
   - Real-time prediction models

3. **External validation**
   - Multi-center studies
   - Prospective evaluation

## Thank You!

\begin{center}
\Large{\textbf{Questions?}}
\end{center}

\vspace{1cm}

**Contact:**

- Ernest Ceballos
- Júlia Galimany  
- Oriol Galimany

\vspace{0.5cm}

*Electronic Health Records - Activity A3*

*Universitat Rovira i Virgili, January 2026*

# Appendix

## Code Availability

- **Analysis Notebook**: `analysis/Activity-A3.ipynb`
- **Full Report**: `paper/main.pdf`
- **Repository**: GitHub - Blanqui04/EHR-A3-Group-B

## References

1. Singer M, et al. *JAMA*. 2016;315(8):801-810. (Sepsis-3 definitions)
2. Johnson AEW, et al. *Scientific Data*. 2016;3:160035. (MIMIC-III)
3. Fleischmann C, et al. *Am J Respir Crit Care Med*. 2016;193(3):259-272.
